---
title: "Melbourne Housing Market-Paper1"
subtitle: "Intro Data Science"
author: "Team 2"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_section: true

---
# Introduction
This paper discusses our project using data on housing prices in Melbourne in 2017.
The establishment of a real estate price prediction system is a key task for the healthy development of the current real estate industry. 
Having a simple predictive and inferential method to model housing prices helps commerce determine fair prices and allows governments to determine property taxes. 
This project aims to learn how different factors may affect home sales price by building linear models. 
Although data that will be utilized was collected in Melbourne, Australia in 2017, the concept that location and home attributes correlate with housing prices could reasonably apply broadly and internationally. 

The following questions are the main subjects which this project focuses on:

1. Understand if housing prices in Melbourne, Australia can be predicted using this dataset.
2. Determine what variables have the greatest impact on housing price.
3. Analyze the impacts of location, seller, and construction attributes of homes on the housing market in Melbourne, Australia. 



```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = F)
options(scientific=T, digits = 5) 
options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times

loadPkg("leaps")
loadPkg("ggplot2")
loadPkg("gridExtra")
loadPkg("scales")
loadPkg("heatmaply")
loadPkg("Hmisc")
loadPkg("faraway")
loadPkg("jtools")

set.seed(17)
data.full <- read.csv("melb_data.csv")

# create train and test data
samps <- read.csv("sample.txt", sep=" ")

data.train <- data.full[(samps$x),]
data.test <- data.full[(-samps$x),]

```

```{r xkablesummary}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")

xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' Can also use as head for better display
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  #' xkabledply( ISLR::Hitters[1:5,] )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}

xkabledplyhead = function(df, rows=5, title="Head", digits = 4, pos="left", bso="striped") { 
  xkabledply(df[1:rows, ], title, digits, pos, bso, wide=FALSE)
}

xkabledplytail = function(df, rows=5, title="Tail", digits = 4, pos="left", bso="striped") { 
  trows = nrow(df)
  xkabledply(df[ (trows-rows+1) : trows, ], title, digits, pos, bso, wide=FALSE)
}

xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}

# this is wrong
xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=TRUE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters, wide=T ) )

  df <- data.frame(vif(model))                            
  names(df) <- "VIF"
     
  if (wide) { df <- t(df) }
  xkabledply(df, title=title, digits = digits, pos=pos, bso=bso )    
}
```

```{r outlierKD2}       
# Fix outliers
outlierKD2 <- function(df, var, rm=FALSE, boxplt=FALSE, histogram=TRUE, qqplt=FALSE, variable_name=deparse(substitute(var))) { 
    #' Original outlierKD functino by By Klodian Dhana,
    #' https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/
    #' Modified to have third argument for removing outliers instead of interactive prompt, 
    #' and after removing outlier, original df will not be changed. The function returns the a df, 
    #' which can be saved as original df name if desired.
    #' Also added QQ-plot in the output, with options to show/hide boxplot, histogram, qqplot.                                 
    #' Check outliers, and option to remove them, save as a new dataframe. 
    #' @param df The dataframe.                                          
    #' @param var The variable in the dataframe to be checked for outliers
    #' @param rm Boolean. Whether to remove outliers or not.          
    #' @param boxplt Boolean. Whether to show the boxplot, before and after outliers removed.
    #' @param histogram Boolean. Whether to show the histogram, before and after outliers removed.
    #' @param qqplt Boolean. Whether to show the qqplot, before and after outliers removed.                                   
    #' @param variable_name String. Variable name formatted for title.
    #' @return The dataframe with outliers replaced by NA if rm==TRUE, or df if nothing changed               
    #' @examples    
    #' outlierKD2(mydf, height, FALSE, TRUE, TRUE, TRUE)
    #' mydf = outlierKD2(mydf, height, TRUE, TRUE, TRUE, TRUE)
    #' mydfnew = outlierKD2(mydf, height, TRUE)
    dt = df # duplicate the dataframe for potential alteration
    var_name <- eval(substitute(var),eval(dt))     
    na1 <- sum(is.na(var_name))
    m1 <- mean(var_name, na.rm = T)
    par(mfrow=c(2, boxplt+histogram+qqplt), oma=c(0,0,3,0))
    if (qqplt) { 
      qqnorm(var_name, main = "With outliers")
      qqline(var_name)
    }
    if (histogram) { hist(var_name, main="With outliers", xlab=NA, ylab=NA) }
    if (boxplt) { boxplot(var_name, main="With outliers") }

    outlier <- boxplot.stats(var_name)$out
    mo <- mean(outlier)
    var_name <- ifelse(var_name %in% outlier, NA, var_name)
    if (qqplt) { 
      qqnorm(var_name, main = "Without outliers")
      qqline(var_name)
    }
    if (histogram) { hist(var_name, main="Without outliers", xlab=NA, ylab=NA) }
    if (boxplt) { boxplot(var_name, main="Without outliers") }

    title(paste("Outlier Check:", variable_name), outer=TRUE)
    na2 <- sum(is.na(var_name))
    cat("Outliers identified:", na2 - na1, "\n")
    cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "\n")
    cat("Mean of the outliers:", round(mo, 2), "\n")
    m2 <- mean(var_name, na.rm = T)
    cat("Mean without removing outliers:", round(m1, 2), "\n")
    cat("Mean if we remove outliers:", round(m2, 2), "\n")
    
    # response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
    # if(response == "y" | response == "yes"){
    if(rm){
        dt[as.character(substitute(var))] <- invisible(var_name)
        #assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
        cat("Outliers successfully removed", "\n")
        return(invisible(dt))
    } else {
        cat("Nothing changed", "\n")
        return(invisible(df))
    }
}
```

```{r create_ggplot_labes}
# factor level counts for legend
counts <- function(variable){                      
                                  
  table.temp <- table(variable)
  regions <- c(rep(0, length(table.temp)))                                            
  
  for (i in 1:length(table.temp)){
    name.new <- paste(names(table.temp[i]), table.temp[[i]], sep=" - ")
    regions[i] <- name.new
}
                                     
  return(regions)
}

```



## The Melbourne Housing Snapshot Dataset

```{r data_feats}                                                                                                                                                                                 

nvar <- ncol(data.full)
nObs <- nrow(data.full)

numerics <- c("Rooms", "Price", "Distance", "Bedroom2", "Bathroom", 
              "Car", "Landsize", "BuildingArea", "YearBuilt", 
              "Lattitude", "Longtitude",
              "Propertycount")                             

numerics.fixed <- c("Rooms", "Price", "Distance", "Bedrooms", "Bathrooms", 
              "Car Parks", "Land size", "Building Area", "Year Built", 
              "Lattitude", "Longtitude",
              "Property Count")
                       
data.fixedname <- data.full                                     
colnames(data.fixedname)[colnames(data.fixedname) %in% numerics] <- numerics.fixed

```

- Home Sales in 2017
  - Location                              
  - Construction
  - Sale
- Variables: `r nvar`
  - Numeric: `r length(numerics)`
  - Categorical: `r nvar - length(numerics)`
                                                           
## The Variables
```{r vars, results="asis"}
var.names <- c(names(data.full))

```

Rooms: Number of rooms
                                                                                          
Price: Price (AUS$)

Method: Method of sale - 5 categories      

Type:  House, Unit, Townhouse - 3 categories   

SellerG: Real Estate Agent - 268 categories             

Date: Date sold                                                                                 

Distance: Distance from Central Business District

Regionname: Region name - 8 categories

Propertycount: Number of properties that exist in the suburb

## More Variables
                                                                    
Bedroom2 : Number of Bedrooms                              
                                                        
Bathroom: Number of Bathrooms                                                                                                    

Car: Number of carspots                                                                                                   

Landsize: Land Size                             

BuildingArea: Building Size

YearBuilt: Year home built

CouncilArea: Governing council for the area - 34 categories

Lattitude, Longtitude: GPS location

Suburb: Suburb name - 314 categories                                            

## Goals
1.	Understand which attributes of a home and its sale determine final sale price
2.	Attempt to build a reasonable model for inference and/or prediction for final sale price

# Exploratory Data Analysis (EDA)
The following are excerpts and graphs from our exploratory data analysis. 
This part of the project familiarizes the reader with our dataset's attributes as well as lays
the foundation for the variables will include in our linear model. 
The results of our EDA will also inform the future direction of the project.

## Summary of Price Statistics

Mean: `r dollar(mean(data.full$Price))`

SD: `r sqrt(var(data.full$Price))`

`r xkablesummary( as.data.frame(data.full$Price), title=NULL)`

## Select Data Pairs
```{r pairs}
pairs(data.full[numerics[c(2, 3, 7, 8, 12)]])
```

## Corrleations

```{r heatmap}

data.corr <- rcorr(as.matrix(data.fixedname[numerics.fixed]))

hm <- heatmaply_cor(
  data.corr$r,
  main="Correlation Heatmap for Numeric Variables",
  custom_hovertext=matrix(paste("p-value:", round(data.corr$P, digits=4)), nrow=length(numerics.fixed)),
  dendrogram="none",
  grid_color="black",
  label_names = c("x", "y", "Correlation")
)

```
`r hm`

## Map of Melbourne  Sales

```{r price_region_qq}


ggplot(data.full) + 
  aes(x=Lattitude, y=Longtitude, color=Regionname) + 
  geom_point()

```


## Selling Price

```{r price_data_plots}


layout(matrix(c(1,2,3), 1,3))
hist(data.full$Price, col = "blue", xlab="Price (AUS$)", main=NULL)
boxplot(data.full$Price, col = "red", ylab="Price (AUS$)")
qqnorm(data.full$Price, col = "dark green", ylab="Price (AUS$)")
qqline(data.full$Price)

```


## Log Selling Price

```{r log_price_plots}
layout(matrix(c(1,2,3), 1,3))
hist(log(data.full$Price), col = "blue", xlab="log(Price)", main=NULL)
boxplot(log(data.full$Price), col = "red", ylab="log(Price)")
qqnorm(log(data.full$Price), col = "dark green", ylab="log(Price)")
qqline(log(data.full$Price))

logPrice.ttest <- t.test(log(data.full$Price), conf.level=0.95)

Price.lower <- exp(logPrice.ttest$conf.int[1])
Price.upper <- exp(logPrice.ttest$conf.int[2])

```


## Price by Region

```{r price_region}

region.count <- counts(data.full$Regionname)

ggplot(data.full, 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + 
  labs(y="Price (AUS$)") + 
  scale_fill_discrete(name="Region Name", labels=region.count) +
  scale_y_continuous(labels = scales::comma)

```






## Price by Number of Rooms (<10 Rooms)

```{r price_rooms}

rooms.count <- counts(data.full$Rooms)

ggplot(subset(data.full, Rooms<10), 
       aes(x=factor(Rooms), y=Price, fill=factor(Rooms))) + 
  geom_boxplot() + 
  labs(x="Number of Rooms", y="Price (AUS$)") + 
  scale_y_continuous(labels = scales::comma) +
  scale_fill_discrete(name="Number of Rooms - Obs", labels=rooms.count) 

```


## Price by Type of Home

```{r price_type}

type.count <- counts(data.full$Type)

ggplot(data.full, 
       aes(x=factor(Type), y=Price, fill=Type)) + 
  geom_boxplot() + 
  labs(y="Price (AUS$)") + 
  scale_y_continuous(labels = scales::comma) +
  guides(fill=FALSE) + 
  scale_x_discrete(name="Type of Home", labels=c("House", "Townhouse", "Apartment"))

```


## Test of Independence (Pearson $\chi^2$) by Group: Type, Rooms, Regionname, SellerG


```{r type_person}

data.full$Price.cuts <- as.character(cut(data.full$Price, 10))
data.full$Price.cuts[data.full$Price > 3650000] <- "(3.65e+06,9.01e+06]"

data.full$Price.cuts <- as.factor(data.full$Price.cuts)


type.chi <- chisq.test(data.full$Type, data.full$Price.cut)
rooms.chi <- chisq.test(subset(data.full, data.full$Rooms<10)$Rooms, subset(data.full, data.full$Rooms<10)$Price.cut)
region.chi <- chisq.test(data.full$Regionname, data.full$Price.cut)

seller.chi <- chisq.test(data.full$SellerG, data.full$Price.cut)
```

$H_0$: All means equal by group 


All reject $H_0$ with p-value$<2\times 10^{-16}$



## Price by Region and Type


```{r price_type_region}

p1 <- ggplot(subset(data.full, Type=="h"), 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() + 
  labs(title="Houses", y="Price (AUS$)") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(labels = scales::comma) +
  guides(fill=FALSE)




p2<- ggplot(subset(data.full, Type=="t"), 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() + 
  labs(title="Townhouses", y=NULL) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(labels = scales::comma) +
  guides(fill=FALSE)



p3 <- ggplot(subset(data.full, Type=="u"), 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() + 
  labs(title="Apartments", y=NULL) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  scale_y_continuous(labels = scales::comma) +
  guides(fill=FALSE)

grid.arrange(p1,p2,p3, nrow=1)

```


# Linear Modelling

## First Attempt at Linear Model

```{r lm1, results='markdown'}

mod <- lm(Price ~ Rooms + Landsize
          + Distance + Bedroom2 + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train)

mod.vifs <- vif(mod)

plot(mod.vifs,
     xlab="", ylab="VIF",
     xaxt='n',
     main="VIF Values for First Model")
abline(h=5, col="blue")

points(which.max(mod.vifs), max(mod.vifs), pch=23, col="red", bg="red")
text(which.max(mod.vifs), max(mod.vifs)-1, label=names(mod$coefficients[which.max(mod.vifs)+1]))

```


## Linear Model 2: Removed the Variable with Highest VIF
```{r lm2_vif, results='markdown'}
# drop Bedroom2 - keep
mod2 <-  lm(Price ~ Rooms + Landsize
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train)


mod2.vifs <- vif(mod2)

plot(mod2.vifs,
     xlab="", ylab="VIF",
     xaxt='n',
     main="VIF Values for Second Model")
abline(h=5, col="blue")

points(which.max(mod2.vifs), max(mod2.vifs), pch=23, col="red", bg="red")
text(which.max(mod2.vifs)-3, max(mod2.vifs)-0.5, label=names(mod2$coefficients[which.max(mod2.vifs)+1]))

```

### Model Coefficients

```{r lm2_mod, results='markdown'}

mod2.summ <-summary(mod2)
plot(mod2.summ$coefficients[,"Pr(>|t|)"],
     xlab="", ylab="p-value",
     xaxt='n',                 
     main="p-values for Second Model")
abline(h=0.05, col="blue")

points(11:mod2$rank, mod2.summ$coefficients[,"Pr(>|t|)"][11:mod2$rank], pch=21, col="blue", bg="blue")
points(3,mod2.summ$coefficients[,"Pr(>|t|)"][3], pch=23, col="red", bg="red")
text(3, mod2.summ$coefficients[,"Pr(>|t|)"][3]-0.01, label=names(mod2$coefficients[3]))

legend("topright",
       pch=21,
       pt.bg="blue",
       col="blue",
       legend="Region Factor Levels")
```

## Linear Model 3: Considered Interactions

```{r lm3, results='markdown'}
#  mod 2 with rooms*region interaction -  failed model, drop interaction
mod3 <-  lm(Price ~ Rooms + Landsize
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname)*Rooms , data=data.train)


mod3.summ <-summary(mod3)                                                                                          
plot(mod3.summ$coefficients[,"Pr(>|t|)"],
     xlab="", ylab="p-value",                                                                                              
     xaxt='n',
     main="p-values for Model with Interaction")
abline(h=0.05, col="blue")

points(11:17, mod3.summ$coefficients[,"Pr(>|t|)"][11:17], pch=21, col="blue", bg="blue")
points(18:24, mod3.summ$coefficients[,"Pr(>|t|)"][18:24], pch=21, col="darkgreen", bg="darkgreen")

legend("topright",
       pch=21,
       pt.bg=c("blue", "darkgreen"),
       col=c("blue", "darkgreen"),
       legend=c("Region Factor Levels", "Interaction Rooms:Region"))



```


## Linear Model 4: Removed Land Size

```{r lm4, results='asis'}

mod4 <-  lm(Price ~ Rooms                                                                                
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train)



mod4.vifs <- vif(mod4)

plot(mod4.vifs,                                               
     xlab="", ylab="VIF",
     xaxt='n',
     main="VIF Values for Forth Model")
abline(h=5, col="blue")

points(which.max(mod4.vifs), max(mod4.vifs), pch=23, col="red", bg="red")
text(which.max(mod4.vifs)-3, max(mod4.vifs)-0.5, label=names(mod4$coefficients[which.max(mod4.vifs)+1]))
```


## Model 4 Coefficients
```{r lm4_pvalues, results='markdown'}

mod4.summ <- summary(mod4)

plot(mod4.summ$coefficients[,"Pr(>|t|)"],
     xlab="", ylab="p-value",
     xaxt='n',
     main="p-values for Second Model")
abline(h=0.05, col="blue")                                                  

points(10:mod4$rank, mod4.summ$coefficients[,"Pr(>|t|)"][10:mod4$rank], pch=21, col="blue", bg="blue")


legend("topright",
       pch=21,
       pt.bg="blue",
       col="blue",
       legend="Region Factor Levels")


```
# Residual Analysis

## Homogeneity? No                                                              
```{r lm4_redids}

plot(mod4, which=1, sub.caption="Model 4")
```

## Normal? Not Quite
```{r lm4_qq}

plot(mod4, which=2, sub.caption="Model 4")                                                                                                                                                      
```

## Influence? Yes
```{r lm4_lev}

plot(mod4, which=5, sub.caption="Model 4")
```



## Remove Influence Points
```{r lm5}
                               
# drop drop influence points
mod5 <- mod4 <-  lm(Price ~ Rooms
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname), data=data.train[!(rownames(data.train) %in% c("13246","2561", "1589")),])

plot(mod4, which=5, sub.caption="Model 4 without Outliers")
```

## Transformed Data - Homogeneity                                                    

```{r lm5_trans}
# log transform
# drop drop influence points
mod6 <- lm(log(Price) ~ Rooms + Car
          + Distance   + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train[!(rownames(data.train) %in% c("13246","2561", "1589")),])

plot(mod6, which=1, sub.caption="Model 4 with Log(Price) Transformation")


y.trans.pred<-predict(mod6)

```

## Transformed Data - Normality                                                               
```{r lm5_trans_qq}
# log transform
# drop drop influence points
mod6 <- lm(log(Price) ~ Rooms + Car
          + Distance   + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train[!(rownames(data.train) %in% c("13246","2561", "1589")),])

plot(mod6, which=2, sub.caption="Model 4 with Log(Price) Transformation")
```

                                                                                                       
# Proposed Model 

``` {r prop_mod, results="markdown"}

summ(mod4, vifs=T)
```



## Testing $R^2$

```{r lm4_pred}
y.pred <- predict(mod4, data.test)                                        
y.test <- data.test$Price                          

diff.sq = (y.pred - y.test)^2 

keeps <- which(!is.na(diff.sq))

r2 <- function(y.predict, y.actual=y.test){
  
  TSS <- sum((y.actual - mean(y.actual))^2)
  RSS <- sum((y.predict - y.actual)^2)
  
  rSq <- 1 - RSS/TSS
  
  return(rSq)
}
    
r.2 <- r2(y.pred[keeps], y.test[keeps])                                           

```

$$ \begin{equation} R^2 = 1- \dfrac{RSS}{TSS} \end{equation}=0.441$$ 



# Conclusion

## Future Work
- Further explore log transformation                 
- Consider GLM with log link
- What to do about factors with many levels (100's)? 
- Missing data
- Improve Prediction

# Bibliography 


