---
title: "Melbourne Housing Market"
subtitle: "Intro Data Science"
author: "Team 2"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r basic, include=F}
# use this function to conveniently load libraries and work smoothly with knitting
# can add quietly=T option to the require() function
# the loadPkg function essentially replaced/substituted two functions install.packages() and library() in one step.
loadPkg = function(x) { if (!require(x,character.only=T, quietly =T)) { install.packages(x,dep=T,repos="http://cran.us.r-project.org"); if(!require(x,character.only=T)) stop("Package not found") } }

# unload/detact package when done using it
unloadPkg = function(pkg, character.only = FALSE) { 
  if(!character.only) { pkg <- as.character(substitute(pkg)) } 
  search_item <- paste("package", pkg,sep = ":") 
  while(search_item %in% search()) { detach(search_item, unload = TRUE, character.only = TRUE) } 
}
```

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# knitr::opts_chunk$set(include = F)
knitr::opts_chunk$set(echo = F)
options(scientific=T, digits = 5) 
options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times

loadPkg("leaps")
loadPkg("ggplot2")
loadPkg("gridExtra")
loadPkg("scales")
loadPkg("heatmaply")
loadPkg("Hmisc")
loadPkg("faraway")

set.seed(17)
data.full <- read.csv("melb_data.csv")

# create train and test data
samps <- sample(nrow(data.full), nrow(data.full)*0.7)

data.train <- data.full[(samps),]
data.test <- data.full[(-samps),]

```

```{r xkablesummary}
loadPkg("xtable")
loadPkg("kableExtra")
loadPkg("stringi")

xkabledply = function(modelsmmrytable, title="Table", digits = 4, pos="left", bso="striped", wide=FALSE) { 
  #' Combining base::summary, xtable, and kableExtra, to easily display model summary. 
  #' wrapper for the base::summary function on model objects
  #' Can also use as head for better display
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param modelsmmrytable This can be a generic table, a model object such as lm(), or the summary of a model object summary(lm()) 
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return HTML table for display
  #' @examples
  #' library("xtable")
  #' library("kableExtra")
  #' xkabledply( df, title="Table testing", pos="left", bso="hover" )
  #' xkabledply( ISLR::Hitters[1:5,] )
  if (wide) { modelsmmrytable <- t(modelsmmrytable) }
  modelsmmrytable %>%
    xtable() %>% 
    kable(caption = title, digits = digits) %>%
    kable_styling(bootstrap_options = bso, full_width = FALSE, position = pos)
}

xkabledplyhead = function(df, rows=5, title="Head", digits = 4, pos="left", bso="striped") { 
  xkabledply(df[1:rows, ], title, digits, pos, bso, wide=FALSE)
}

xkabledplytail = function(df, rows=5, title="Tail", digits = 4, pos="left", bso="striped") { 
  trows = nrow(df)
  xkabledply(df[ (trows-rows+1) : trows, ], title, digits, pos, bso, wide=FALSE)
}

xkablesummary = function(df, title="Table: Statistics summary.", digits = 4, pos="left", bso="striped") { 
  #' Combining base::summary, xtable, and kableExtra, to easily display numeric variable summary of dataframes. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param df The dataframe.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @return The HTML summary table for display, or for knitr to process into other formats 
  #' @examples
  #' xkablesummary( faraway::ozone )
  #' xkablesummary( ISLR::Hitters, title="Five number summary", pos="left", bso="hover"  )
  
  s = summary(df) %>%
    apply( 2, function(x) stringr::str_remove_all(x,c("Min.\\s*:\\s*","1st Qu.\\s*:\\s*","Median\\s*:\\s*","Mean\\s*:\\s*","3rd Qu.\\s*:\\s*","Max.\\s*:\\s*")) ) %>% # replace all leading words
    apply( 2, function(x) stringr::str_trim(x, "right")) # trim trailing spaces left
  
  colnames(s) <- stringr::str_trim(colnames(s))
  
  if ( dim(s)[1] ==6 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max') 
  } else if ( dim(s)[1] ==7 ) { rownames(s) <- c('Min','Q1','Median','Mean','Q3','Max','NA') }
  
  xkabledply(s, title=title, digits = digits, pos=pos, bso=bso )
}

# this is wrong
xkablevif = function(model, title="VIFs of the model", digits = 3, pos="left", bso="striped", wide=TRUE) { 
  #' Combining faraway::vif, xtable, and kableExtra, to easily display numeric summary of VIFs for a model. 
  #' ELo 202004 GWU DATS
  #' version 1.2
  #' @param model The lm or compatible model object.
  #' @param title Title of table. 
  #' @param digits Number of digits to display
  #' @param pos Position of table, c("left","center","right") 
  #' @param bso bootstrap_options = c("basic", "striped", "bordered", "hover", "condensed", "responsive")
  #' @param wide print table in long (FALSE) format or wide (TRUE) format
  #' @return The HTML summary table of the VIFs for a model for display, or for knitr to process into other formats 
  #' @examples
  #' xkablevif( lm(Salary~Hits+RBI, data=ISLR::Hitters, wide=T ) )

  df <- data.frame(vif(model))
  names(df) <- "VIF"
  
  if (wide) { df <- t(df) }
  xkabledply(df, title=title, digits = digits, pos=pos, bso=bso )
}
```

```{r outlierKD2}
# Fix outliers
outlierKD2 <- function(df, var, rm=FALSE, boxplt=FALSE, histogram=TRUE, qqplt=FALSE, variable_name=deparse(substitute(var))) { 
    #' Original outlierKD functino by By Klodian Dhana,
    #' https://www.r-bloggers.com/identify-describe-plot-and-remove-the-outliers-from-the-dataset/
    #' Modified to have third argument for removing outliers instead of interactive prompt, 
    #' and after removing outlier, original df will not be changed. The function returns the a df, 
    #' which can be saved as original df name if desired.
    #' Also added QQ-plot in the output, with options to show/hide boxplot, histogram, qqplot.
    #' Check outliers, and option to remove them, save as a new dataframe. 
    #' @param df The dataframe.
    #' @param var The variable in the dataframe to be checked for outliers
    #' @param rm Boolean. Whether to remove outliers or not.
    #' @param boxplt Boolean. Whether to show the boxplot, before and after outliers removed.
    #' @param histogram Boolean. Whether to show the histogram, before and after outliers removed.
    #' @param qqplt Boolean. Whether to show the qqplot, before and after outliers removed.
    #' @param variable_name String. Variable name formatted for title.
    #' @return The dataframe with outliers replaced by NA if rm==TRUE, or df if nothing changed
    #' @examples
    #' outlierKD2(mydf, height, FALSE, TRUE, TRUE, TRUE)
    #' mydf = outlierKD2(mydf, height, TRUE, TRUE, TRUE, TRUE)
    #' mydfnew = outlierKD2(mydf, height, TRUE)
    dt = df # duplicate the dataframe for potential alteration
    var_name <- eval(substitute(var),eval(dt))
    na1 <- sum(is.na(var_name))
    m1 <- mean(var_name, na.rm = T)
    par(mfrow=c(2, boxplt+histogram+qqplt), oma=c(0,0,3,0))
    if (qqplt) { 
      qqnorm(var_name, main = "With outliers")
      qqline(var_name)
    }
    if (histogram) { hist(var_name, main="With outliers", xlab=NA, ylab=NA) }
    if (boxplt) { boxplot(var_name, main="With outliers") }

    outlier <- boxplot.stats(var_name)$out
    mo <- mean(outlier)
    var_name <- ifelse(var_name %in% outlier, NA, var_name)
    if (qqplt) { 
      qqnorm(var_name, main = "Without outliers")
      qqline(var_name)
    }
    if (histogram) { hist(var_name, main="Without outliers", xlab=NA, ylab=NA) }
    if (boxplt) { boxplot(var_name, main="Without outliers") }

    title(paste("Outlier Check:", variable_name), outer=TRUE)
    na2 <- sum(is.na(var_name))
    cat("Outliers identified:", na2 - na1, "\n")
    cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "\n")
    cat("Mean of the outliers:", round(mo, 2), "\n")
    m2 <- mean(var_name, na.rm = T)
    cat("Mean without removing outliers:", round(m1, 2), "\n")
    cat("Mean if we remove outliers:", round(m2, 2), "\n")
    
    # response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
    # if(response == "y" | response == "yes"){
    if(rm){
        dt[as.character(substitute(var))] <- invisible(var_name)
        #assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
        cat("Outliers successfully removed", "\n")
        return(invisible(dt))
    } else {
        cat("Nothing changed", "\n")
        return(invisible(df))
    }
}
```

```{r create_ggplot_labes}
# factor level counts for legend
counts <- function(variable){
  
  table.temp <- table(variable)
  regions <- c(rep(0, length(table.temp)))
  
  for (i in 1:length(table.temp)){
    name.new <- paste(names(table.temp[i]), table.temp[[i]], sep=" - ")
    regions[i] <- name.new
}
  
  return(regions)
}

```



## The Melbourne Housing Snapshot Dataset

```{r data_feats}

nvar <- ncol(data.full)
nObs <- nrow(data.full)

numerics <- c("Rooms", "Price", "Distance", "Bedroom2", "Bathroom", 
              "Car", "Landsize", "BuildingArea", "YearBuilt", 
              "Lattitude", "Longtitude",
              "Propertycount")

numerics.fixed <- c("Rooms", "Price", "Distance", "Bedrooms", "Bathrooms", 
              "Car Parks", "Land size", "Building Area", "Year Built", 
              "Lattitude", "Longtitude",
              "Property Count")

data.fixedname <- data.full
colnames(data.fixedname)[colnames(data.fixedname) %in% numerics] <- numerics.fixed

```

- Home Sales in 2017
  - Location
  - Construction
  - Sale
- Variables: `r nvar`
  - Numeric: `r length(numerics)`
  - Categorical: `r nvar - length(numerics)`


## Summary of Select Numeric Variables

```{r price_summary}

summ <- summary(data.full[c("Price", "Rooms", "BuildingArea")])

```

`r xkablesummary(data.full[c("Price", "Rooms", "BuildingArea")], title=NULL)`

## Select Data Pairs
```{r pairs}
pairs(data.full[numerics[c(2, 3, 7, 8, 12)]])
```

## Corrleations

```{r heatmap}

data.corr <- rcorr(as.matrix(data.fixedname[numerics.fixed]))

hm <- heatmaply_cor(
  data.corr$r,
  main="Correlation Heatmap for Numeric Variables",
  custom_hovertext=matrix(paste("p-value:", round(data.corr$P, digits=4)), nrow=length(numerics.fixed)),
  dendrogram="none",
  grid_color="black",
  label_names = c("x", "y", "Correlation")
)

```
`r hm`

## Map of Melbourne Sales

```{r price_region_qq}


ggplot(data.full) + 
  aes(x=Lattitude, y=Longtitude, color=Regionname) + 
  geom_point()

```


## Selling Price

```{r price_data_plots}


# 
# p1 <- ggplot(data.full, aes(x=Price)) + 
#   geom_histogram(fill="blue", color="black") +
#   labs(x="Price (AUS$)", y="Frequency") + 
#   scale_x_continuous(labels = scales::comma) +
#   theme(aspect.ratio=1)
# 
# p2 <- ggplot(data.full, aes(y=Price)) + 
#   geom_boxplot(fill="blue") +
#   labs(y="Price (AUS$)") + 
#   scale_y_continuous(labels = scales::comma) +
#   theme(axis.title.x=element_blank(),
#         axis.text.x=element_blank(),
#         axis.ticks.x=element_blank(),
#         aspect.ratio=1) 
# 
# p3 <- ggplot(data.full, aes(sample=Price)) + 
#   stat_qq(color="blue") + 
#   stat_qq_line() + 
#   labs(y="Price (AUS$)", x="Theoretical") +
#   scale_y_continuous(labels = scales::comma) +
#   theme(aspect.ratio=1)
# 
# 
# grid.arrange(p1, p2, p3, nrow=1)


layout(matrix(c(1,2,3), 1,3))
hist(data.full$Price, col = "blue", xlab="Price (AUS$)", main=NULL)
boxplot(data.full$Price, col = "red", ylab="Price (AUS$)")
qqnorm(data.full$Price, col = "dark green", ylab="Price (AUS$)")
qqline(data.full$Price)

```


## Log Selling Price

```{r log_price_plots}
layout(matrix(c(1,2,3), 1,3))
hist(log(data.full$Price), col = "blue", xlab="log(Price)", main=NULL)
boxplot(log(data.full$Price), col = "red", ylab="log(Price)")
qqnorm(log(data.full$Price), col = "dark green", ylab="log(Price)")
qqline(log(data.full$Price))

logPrice.ttest <- t.test(log(data.full$Price), conf.level=0.95)

Price.lower <- exp(logPrice.ttest$conf.int[1])
Price.upper <- exp(logPrice.ttest$conf.int[2])

```


## Price by Region

```{r price_region}

region.count <- counts(data.full$Regionname)

ggplot(data.full, 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + 
  labs(y="Price (AUS$)") + 
  scale_fill_discrete(name="Region Name", labels=region.count) +
  scale_y_continuous(labels = scales::comma)

```






## Price by Number of Rooms

```{r price_rooms}

rooms.count <- counts(data.full$Rooms)

ggplot(subset(data.full, Rooms<10), 
       aes(x=factor(Rooms), y=Price, fill=factor(Rooms))) + 
  geom_boxplot() + 
  labs(x="Number of Rooms", y="Price (AUS$)") + 
  scale_y_continuous(labels = scales::comma) +
  scale_fill_discrete(name="Number of Rooms - Obs", labels=rooms.count) 

```


## Price by Type of Home

```{r price_type}

type.count <- counts(data.full$Type)

ggplot(data.full, 
       aes(x=factor(Type), y=Price, fill=Type)) + 
  geom_boxplot() + 
  labs(y="Price (AUS$)") + 
  scale_y_continuous(labels = scales::comma) +
  guides(fill=FALSE) + 
  scale_x_discrete(name="Type of Home", labels=c("House", "Townhouse", "Apartment"))

```



## Price by Region and Type

if we keep this, consider making the legend colors standardized so region 
same no matter the chart

```{r price_type_region}

p1 <- ggplot(subset(data.full, Type=="h"), 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() + 
  labs(title="Houses", y="Price (AUS$)") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(labels = scales::comma) +
  guides(fill=FALSE)




p2<- ggplot(subset(data.full, Type=="t"), 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() + 
  labs(title="Townhouses", y=NULL) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+
  scale_y_continuous(labels = scales::comma) +
  guides(fill=FALSE)



p3 <- ggplot(subset(data.full, Type=="u"), 
       aes(x=Regionname, y=Price, fill=Regionname)) + 
  geom_boxplot() + 
  labs(title="Apartments", y=NULL) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  scale_y_continuous(labels = scales::comma)+
  guides(fill=FALSE)

grid.arrange(p1,p2,p3, nrow=1)

```

## First Attempt at Linear Model

```{r lm1, results='asis'}

mod <- lm(Price ~ Rooms + Landsize
          + Distance + Bedroom2 + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train)

mod.vifs <- vif(mod)

# ggplot(data.frame(mod.vifs), aes(x=1:length(mod.vifs), y=mod.vifs)) +
#   geom_point() +
#   labs(title="VIF Values for First Model",
#        y="VIF") + 
#  scale_x_discrete(name="Variable", labels=names(mod.vifs))
# 

plot(mod.vifs,
     xlab="", ylab="VIF",
     xaxt='n',
     main="VIF Values for First Model")
abline(h=5, col="blue")

points(which.max(mod.vifs), max(mod.vifs), pch=23, col="red", bg="red")
text(which.max(mod.vifs), max(mod.vifs)-1, label=names(mod$coefficients[which.max(mod.vifs)+1]))

```




## Remove the Variable with Highest VIF
```{r lm2_vif, results='asis'}
# drop Bedroom2 - keep
mod2 <-  lm(Price ~ Rooms + Landsize
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train)


mod2.vifs <- vif(mod2)

plot(mod2.vifs,
     xlab="", ylab="VIF",
     xaxt='n',
     main="VIF Values for Second Model")
abline(h=5, col="blue")

points(which.max(mod2.vifs), max(mod2.vifs), pch=23, col="red", bg="red")
text(which.max(mod2.vifs)-3, max(mod2.vifs)-0.5, label=names(mod2$coefficients[which.max(mod2.vifs)+1]))

```

## Model Coefficients

```{r lm2_mod, results='asis'}

mod2.summ <-summary(mod2)
plot(mod2.summ$coefficients[,"Pr(>|t|)"],
     xlab="", ylab="p-value",
     xaxt='n',
     main="p-values for Second Model")
abline(h=0.05, col="blue")

points(11:mod2$rank, mod2.summ$coefficients[,"Pr(>|t|)"][11:mod2$rank], pch=21, col="blue", bg="blue")
points(3,mod2.summ$coefficients[,"Pr(>|t|)"][3], pch=23, col="red", bg="red")
text(3, mod2.summ$coefficients[,"Pr(>|t|)"][3]-0.01, label=names(mod2$coefficients[3]))

legend("topright",
       pch=21,
       pt.bg="blue",
       col="blue",
       legend="Region Factor Levels")
```

## Consider Interactions

```{r lm3, results='asis'}
#  mod 2 with rooms*region interaction -  failed model, drop interaction
mod3 <-  lm(Price ~ Rooms + Landsize
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname)*Rooms , data=data.train)


mod3.summ <-summary(mod3)
plot(mod3.summ$coefficients[,"Pr(>|t|)"],
     xlab="", ylab="p-value",
     xaxt='n',
     main="p-values for Model with Interaction")
abline(h=0.05, col="blue")

points(11:17, mod3.summ$coefficients[,"Pr(>|t|)"][11:17], pch=21, col="blue", bg="blue")
points(18:24, mod3.summ$coefficients[,"Pr(>|t|)"][18:24], pch=21, col="darkgreen", bg="darkgreen")

legend("topright",
       pch=21,
       pt.bg=c("blue", "darkgreen"),
       col=c("blue", "darkgreen"),
       legend=c("Region Factor Levels", "Interaction Rooms:Region"))



```


## Remove Land Size

```{r lm4, results='asis'}

mod4 <-  lm(Price ~ Rooms
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train)



mod4.vifs <- vif(mod4)

plot(mod4.vifs,
     xlab="", ylab="VIF",
     xaxt='n',
     main="VIF Values for Forth Model")
abline(h=5, col="blue")

points(which.max(mod4.vifs), max(mod4.vifs), pch=23, col="red", bg="red")
text(which.max(mod4.vifs)-3, max(mod4.vifs)-0.5, label=names(mod4$coefficients[which.max(mod4.vifs)+1]))
```


## Model 4 Coefficients
```{r lm4_pvalues, results='asis'}

mod4.summ <- summary(mod4)

plot(mod4.summ$coefficients[,"Pr(>|t|)"],
     xlab="", ylab="p-value",
     xaxt='n',
     main="p-values for Second Model")
abline(h=0.05, col="blue")

points(10:mod4$rank, mod4.summ$coefficients[,"Pr(>|t|)"][10:mod4$rank], pch=21, col="blue", bg="blue")


legend("topright",
       pch=21,
       pt.bg="blue",
       col="blue",
       legend="Region Factor Levels")


```

## Resdidual Analysis - Homogeneity? No
```{r lm4_redids}

plot(mod4, which=1, sub.caption="Model 4")
```

## Resdidual Analysis - Normal? Not Quite
```{r lm4_qq}

plot(mod4, which=2, sub.caption="Model 4")
```

## Resdidual Analysis - Influence? Yes
```{r lm4_lev}

plot(mod4, which=5, sub.caption="Model 4")
```



## Remove Influence Points
```{r lm5}

# drop drop influence points
mod5 <- mod4 <-  lm(Price ~ Rooms
          + Distance  + Bathroom + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname), data=data.train[!(rownames(data.train) %in% c("13246","2561", "1589")),])

plot(mod4, which=5, sub.caption="Model 4 without Outliers")
```


## Testing $R^2$

```{r lm4_pred}
y.pred <- predict(mod4, data.test)
y.test <- data.test$Price

diff.sq = (y.pred - y.test)^2 

keeps <- which(!is.na(diff.sq))

r2 <- function(y.predict, y.actual=y.test){
  
  TSS <- sum((y.actual - mean(y.actual))^2)
  RSS <- sum((y.predict - y.actual)^2)
  
  rSq <- 1 - RSS/TSS
  
  return(rSq)
}

r.2 <- r2(y.pred[keeps], y.test[keeps])

```

$$ \begin{equation} R^2 = 1- \dfrac{RSS}{TSS} \end{equation}=0.441$$ 



## Transform Data - Homogeneity

```{r lm5_trans}
# log transform
# drop drop influence points
mod6 <- lm(log(Price) ~ Rooms + Car
          + Distance   + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train[!(rownames(data.train) %in% c("13246","2561", "1589")),])

plot(mod6, which=1, sub.caption="Model 4 with Log(Price) Transformation")
```

## Transform Data - Normality
```{r lm5_trans_qq}
# log transform
# drop drop influence points
mod6 <- lm(log(Price) ~ Rooms + Car
          + Distance   + Car + BuildingArea
          + Lattitude + Longtitude + Propertycount
          + factor(Regionname) , data=data.train[!(rownames(data.train) %in% c("13246","2561", "1589")),])

plot(mod6, which=2, sub.caption="Model 4 with Log(Price) Transformation")
```


## Future Work
- Further explore log transformation
- Consider GLM
- What to do about factors with many levels (100's)? 
- Missing data
- Improve Prediction

